---
title: "How to Make a Proceedings Paper Submission"
bibliography: library.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA
    \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\ Department of Educational Psychology, 1025 W. Johnson Street \\ Madison, WI 53706 USA}

abstract: >
    Include no author information in the initial submission, to facilitate
    blind review.  The abstract should be one paragraph, indented 1/8 inch on both sides,
    in 9~point font with single spacing. The heading 'Abstract'
    should be 10~point, bold, centered, with one line of space below
    it. This one-paragraph abstract section is required only for standard
    six page proceedings papers. Following the abstract should be a blank
    line, followed by the header 'Keywords' and a list of
    descriptive keywords separated by semicolons, all in 9~point font, as
    shown below.
    
keywords: >
    Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
    
output: cogsci2024::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Intro

when is eavesdropping effective? 

reference games are used as a microcosm for understanding convention formation with claims of partner specificity 

it's unclear how and when these conventions are transparent to others
* cite the old work about side participants etc
* may be able to understand more than what you'd say (ex. regionalisms)
how well can people understand descriptions from reference games, and what accounts for variation (iconicity of targets, when the references came from, structure of interaction)
* possibly introduce ideas of schelling point or "priors"
but depends on what we think we're coming back to in discussion!

* mention Judy's paper that has the yoked not yoked part 

Need to introduce model perspective!
* models are useful in that we can run expts on them far in excess of what we can in humans
* caption type models are pretty good 
* cite prior work on trying to add pragmatics to models

Introduce dataset that we use for this 

How good is "eavesdropper" comprehension and what factors matter to better comprehension? (need to set up whatever the factors are in intro!)

# Human expts on isolated accuracy 

expts 1 & 2 

tg-matcher 1& 2 (what veronica ran in May) 
1 is 2 and 6 player games, rounds 1 and 6 in medium thick (60 participants, 60 items each)
2 is 2 and 6 player games, rounds 1 and 6 in thin and thick (60 participants, 64 items each)

Key questions:
* how accurate
* what are biggest predictors (conditions, round, target) 

Large item level differences, but not a lot else. 

## Model + calibration

modeling approach and selection 

how well can models proxy humans? are there differences? 

tg-matcher 3 (calibration)
61 participants, 64 items each, from a pool of 217 transcripts spanning the models full accuracy range

given this proxy -- what can we think we learn -- for instance about most important parts of the utterances? (rolling window?)

## Human yoked v not yoked expt

tg-matcher 4 (SPR + yoked/unyoked)
196 participants (99 in yoked, 97 in shuffled), each saw all 72 trials from 1 of 10 games. 
games not chosen at random

seeing things in the same order helps
look at item level accuracy differences? 

# Discussion

role of context 

limitations, incuding out of distribution for models 

might want to address language comprehension v inference 

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
